Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-14 qpip0.tex
Fix protocols

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave a 4-round interactive protocol to let a $\BQP$ machine "commit a $XZ$ measurement" to a classical machine, which we describe below while omitting the implementation details.

\begin{protocol}{Mahadev's 4 round commitment}
\label{proto:urmila4}
\Ethan{TODO Make notation consistent with protocols in other sections}

\begin{enumerate}

    \item Verifier $\bbV$ is a classical machine. Prover $\bbP$ is a $\BQP$ machine. 
%    \item Both the verifier and the prover know a $\BQP$ circuit $C$ and an input string $x$.
    \item Both party know the number of qubits $n$ and a security parameter $\lambda=1^{\poly(n)}$.
    \item Verifier has a string $h \in \{0,1\}^n$ that specifies the measurement he wants to make: if $h_i=0$, he wants a $X$ measurement on $i$-th qubit. If $h_i=1$, he wants a $Z$ measurement on $i$-th qubit.  
    \item In the first round, the verifier generate a secret key and  public key pair $(sk,pk)$ from $h$ and $\lambda$ and sent $pk$ to the prover.
    \item In the second round, the prover pick some $n$-qubit quantum state $\rho$ and generate a "commitment" $y$ from $pk$ and $\rho$. He send $y$ to the verifier and keep the leftover quantum state $\sigma$.
    \item In the third round, the verifier send a uniformly random challenge bit $c\in \{0,1\}$. The prover needs to do a "Testing round" if $c=0$. The prover needs to do a "Hadamard round" if $c=1$.
    \item In the fourth round. The prover generates a classical string $a$ from $c$ and $\sigma$ and send to back to the prover.
    \item If $c=0$, the verifier decides whether he is going to reject from $(pk,a,y)$. If $c=1$ the verifier computes an $n$-bit string $z$ from $(sk,h,y,a)$, $z=V_f(sk,h,y,a)$.
\end{enumerate}
\end{protocol}

@@

% \begin{lem}[another triangle inequality for computational distance \hannote{unused}]\label{lem:computational-triangle}
% Let $D_1$, $D_2$ be two distributions such that $D_1=D_3+D_4$ and $D_2=D_5+D_6$, where $D_3$, $D_4$, $D_5$, and $D_6$ are subnormalized distributions.

% If 
% \begin{align}
%     &\norm{D_1 - D_2}_q \leq \eps   \\ 
%     &\norm{D_4 - D_6}_q \leq \eps',  
% \end{align}
% then 
% \begin{align}
%   \norm{ D_3 - D_5}_q \leq \eps+\eps' 
% \end{align}
% \end{lem}
% % \begin{align}
% %     D_1 &\approx_{q,\eps} D_2 \\
% %     D_4 &\approx_{q,\eps'} D_6, 
% % \end{align}
% % then 
% % \begin{align}
% %     D_3 \approx_{q,\eps+\eps'} D_5
% % \end{align}
% \begin{proof}
% % Since
% % $$\norm{D_4 - D_6}_q \leq \eps',$$
% % there exist

% We prove by contradiction. Suppose that
% $$\norm{ D_3 - D_5}_q > \eps+\eps',$$
% then there exist a quantum circuit $A$ of size $poly(n)$ such that 
% $$\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| > \eps+\eps'.$$
% Since $$\norm{D_4 - D_6}_q \leq \eps'$$, we know that 
% $$\L\|\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \leq \eps'.$$
% Since $D_1=D_3+D_4$ and $D_2=D_5+D_6$, we have that
% $$ \Pr[A({D_1})=1]= \Pr[A({D_3})=1]+ \Pr[A({D_4})=1]$$ 
% $$ \Pr[A({D_2})=1]= \Pr[A({D_5})=1]+ \Pr[A({D_6})=1].$$ 
% So 
% \begin{align}
%     &\L\|\Pr[A({D_1})=1]-\Pr[A({D_2})=1]  \R\| \nn \\
%     =&   \L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1] +\Pr[A({D_4})=1]-\Pr[A({D_6})=1]  \R\| \nn \\
%     \geq  &\L\|\Pr[A({D_3})=1]-\Pr[A({D_5})=1]  \R\| - \L\|\Pr[A({D_6})=1]-\Pr[A({D_4})=1]  \R\|  \nn \\
%      > &\eps 
% \end{align}
% This contradicts with the assumption that $\norm{D_1 - D_2}_q \leq \eps $.
% \end{proof}

@@

@@

By combining the commitment Protocol, Protocol~\ref{proto:urmila4}, and the $\QPIP_1$ Protocol, Protocol~\ref{ProtoQPIP1}, we get a naive $\QPIP_0$ Protocol for $\SampBQP$ that has no soundness guarantee.

@@

\begin{protocol}{naive $\QPIP_0$ protocol  for the $\SampBQP$ instance $(D_x)_{x\in\set{0, 1}^*}$}
	\label{proto:qpip0_naive}

	Inputs: \Ethan{Just assume PiSamp has the properties I need including non-adaptive measurements. Maybe make author note for now if needed and add it back later}
	\begin{itemize}
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
		\item Error parameter $\varepsilon\in(0, 1)$
	\end{itemize}

 	Ingrediants:
 	\begin{itemize}
 		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
 		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
 		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$.
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
%  		\item Let $\cVGS$ be the way to verify said ground state as defined in \Cref{AlgGroundStateCheck}. \Ethan{TODO wording needs fix...? How?}
 	\end{itemize}

	Protocol:
	\begin{enumerate}
	\item The verifier choose a $XZ$-measurement $h$ over  $M=\frac{T^{21}}{\varepsilon^{33}}$  copies of $\ket{\psi_{C'(x)}}$ from the distribution emergent from by Step~\ref{step:qpip1-verify} of Protocol~\ref{ProtoQPIP1}.

	\item \label{step:urmila-in-naive} The verifier and prover run Protocol~\ref{proto:urmila4}, with the following parameters: $h$ is chosen from previous step, $n_{commit}=M \, s$ , where $n_{commit}$ is the $n$ for Protocol~\ref{proto:urmila4}, and the honest prover choose $M$ copies of $\ket{\psi_{C'(x)}}$ as $\rho$. We also allow a cheating prover to send $\bot$ in the fourth round of Protocol~\ref{proto:urmila4} if $c=1$ to abort the process.\footnote{This is useless for the cheating prover but useful for proving one of our theorems.} If the verifier receives $\bot$ in the fourth round, he outputs $(\Rej,\bot)$ and terminates this protocol. Let $z_{commit}$ be the output of the called Protocol~\ref{proto:urmila4}.

	\item If $c=1$, the verifier finished the verification step of Protocol~\ref{ProtoQPIP1} by using $z_{commit}$  as the measurement result he would get and outputs $(d,z_{naive})$.

	\end{enumerate}
\end{protocol}

@@

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of  Protocol~\ref{proto:qpip0_naive} ]
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen,

@@

\begin{protocol}{$\QPIP_0$ for $\SampBQP$}
\label{proto:QPIP0samp}

\hannote{fix inputs}
Inputs: 
 Both verifier and prover start with the description of number of qubits $n$,  input string $x$ to the $\SampBQP$ Instance, and a soundness parameter $\varepsilon$.

Protocol:
\begin{enumerate}
%    \item Verifier start with description of quantum circuit $C$, input string $x$

    \item Run $m=O(1/\varepsilon^2)$ copies of protocol~\ref{proto:urmila4} in parallel \Ethan{proto 2? Or 3?}, with following modifications. The copies have separate $\{h\}$, $\{sk\}$, $\{pk\}$, $\{y\}$, $\{c\}$, and $\{a\}$.  The verifier draws $\{h\}$ independently from the distribution specifier by step~\ref{step:qpip1-verify} of Protocol~\ref{ProtoQPIP1} to generate $\{sk\}$ and $\{pk\}$ for all copies on the first round. On the second round, for each copy, the prover choose the state specified by step~\ref{step:qpip1-state-gen} of Protocol~\ref{ProtoQPIP1}  (with  $(C,x,\eps)$ as input) as $\rho$. On the third round, the verifier uniformly randomly pick an $i\in [m]$, do a Hadamard round on $i$-th copy, and has do test rounds on all other copies, i.e. $c_i=1$ and $c_j=0$ for all $j\neq i$.
    \item \label{step:multi-testing}Verifier rejects if the result on any of the test rounds is rejected. 
    \item If all testing rounds are accepted, verifier run the verification procedure of Protocol~\ref{ProtoQPIP1} on the string $z$ decoded from the Hadamard round.
\end{enumerate}
\end{protocol}

>>>>

In her groundbreaking work, Mahadev~\cite{FOCS:Mahadev18a} gave a 4-round $\QPIP_0$ protocol to let a $\BQP$ machine "commit a $XZ$ measurement" to a classical machine, which we describe below while omitting the implementation details.

\begin{protocol}{Mahadev's 4 round commitment protocol $\PiMeasure$}
\label{proto:urmila4}
\Ethan{Still rooms for improvements, but it's not as bad as before}

Inputs:
\begin{itemize}
	\item Common input: Security parameter $1^\lambda$ where $\lambda\in\bbN$.
	\item Prover's input: a state $\rho\in\cB^{\otimes n}$ for the verifier to measure.
	\item Verifier's inputs:
		\begin{itemize}
			\item a string $h \in \{0,1\}^n$ that specifies the measurement he wants to make: if $h_i=0$, he wants a $X$ measurement on $i$-th qubit. If $h_i=1$, he wants a $Z$ measurement on $i$-th qubit.  
			\item a challenge bit $c\in\zo$. The prover needs to do a ``Testing round" if $c=0$. The prover needs to do a ``Hadamard round" if $c=1$.
		\end{itemize}
\end{itemize}

The protocol:
\begin{enumerate}
    \item The verifier generates a public and secret key pair $(pk, sk)$ on inputs $h$ and $1^\lambda$. It sends $pk$ to the prover.
    \item The prover generates a classical ``commitment" $y$ and internal state $\sigma$ on inputs $pk$ and $\rho$. He send $y$ to the verifier.
	\item The verifier sends $c$ to the prover.
    \item The prover generates a classical string $a$ on inputs $c$ and $\sigma$, and sends it to back to the verifier.
    \item If $c=0$, the verifier decides whether he is going to reject on inputs $(pk,a,y)$.
		If $c=1$, the verifier computes an $n$-bit string $z=V_f(sk,h,y,a)$ on inputs $(sk,h,y,a)$.
\end{enumerate}
\end{protocol}

\Ethan{The following modifications are mine}
Two modification have been made from the original protocol.
\begin{itemize}
	\item On the protocol's third message. Instead of randomly picking whether to run what's called a ``testing round" or a ``Hadamard round", we let the verifier decide as one of its inputs.
	\item On the protocol's fourth message, the prover can choose to abort by sending $a=\bot$. This is useless for the cheating prover but useful for proving one of our theorems.
\end{itemize}

@@

@@

\Ethan{TODO fix variable labels and whatnot here}

@@

By combining the commitment Protocol, $\PiMeasure$, and our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, we get a naive $\QPIP_0$ Protocol for $\SampBQP$ that has no soundness guarantee.
In particular, we combine them by replacing the verifier's measurements with a call to $\PiMeasure$. It is possible because $\VSamp$ applies only $X$ and $Z$ measurements.
Furthermore, its measurements are not adaptive, meaning measurement results do not change later measurement basis choices.

@@

\begin{protocol}{naive $\QPIP_0$ protocol $\PiNaive$ for the $\SampBQP$ instance $(D_x)_{x\in\set{0, 1}^*}$}
	\label{proto:qpip0_naive}

	Inputs:
	\begin{itemize}
		\item Security parameter $1^\lambda$ where $\lambda\in\bbN$
		\item Error parameter $\varepsilon\in(0, 1)$
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
		\item Verifier's input: challenge bit $c\in\zo$ as in $\PiMeasure$
	\end{itemize}

	% Ingrediant: Let $\varepsilon'=0.99\varepsilon$

	Protocol:
	\begin{enumerate}
		\item The verifier choose a $XZ$-measurement $h$ from the distribution emergent from by Step~\ref{step:qpip1-verify} of $\PiSamp$. \Ethan{I didn't change the wording here}
			It samples $c\xleftarrow{\$}\zo$ uniformly at random.
		\item The prover prepares $\sigma$ by running Step~\ref{step:qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\sigma), \VMeasure(h))(1^\lambda, 1^{1/\varepsilon}, x)$.
			If $c=1$, the verifier obtains $z_{commit}$.
		\item If $c=1$, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by using $z_{commit}$ as the measurement result he would get and outputs $(d,z_{naive})$.
	\end{enumerate}
\end{protocol}

@@

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of  Protocol~\ref{proto:qpip0_naive} ]
	\Ethan{I haven't done anything with this lemma yet. Might need to fix or make consistent later.}
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen,

@@

\begin{protocol}{$\QPIP_0$ protocol $\PiSampZ$ for $\SampBQP$}
\label{proto:QPIP0samp}

Inputs:
	\begin{itemize}
		\item Security parameter $1^\lambda$ for $\lambda\in\bbN$.
		\item Accuracy paramter $1^\varepsilon$ for the $\SampBQP$ instance
		\item Input $x\in\zo^n$ \Ethan{Or $\zo^*$?} for the $\SampBQP$ instance
	\end{itemize}

Ingrediant: Let $m=O(1/\varepsilon^2)$ be the number of parallel repetitions to run.

Protocol:
\begin{enumerate}
%    \item Verifier start with description of quantum circuit $C$, input string $x$ 
	\item The verifier samples $r\xleftarrow{\$}[m]$. For $1\leq i\leq m$, if $i\ne r$ then set $c_i\leftarrow 0$, else set $c_i\leftarrow 1$.
	\item The prover and verifier runs $m$-fold parallel repetion of $\PiNaive$.
		That is, for $1\leq i\leq m$, run $(\PNaive, \VNaive(c_i))(1^\lambda, 1^\varepsilon, x)$.
	\item \label{step:multi-testing} If any of the testing round copies are rejected, the verifier outputs $(\Rej, \bot)$.
		Else it outputs the result from the Hadamard round copy.
\end{enumerate}
\end{protocol}
